---
title: "Injecting coleo data"
author: "Andrew MacDonald"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r, include=FALSE}
library(httptest2)
start_vignette("coleo_data_injection")
```

## injecting cells

We start with an example of data from an imaginary cell:

```{r dput_cell, include=FALSE}
cell_data <- structure(list(cell_code = "111_777", name = "Middle Earth", 
    geom = structure(list(structure(list(structure(c(-68.784529, 
    -68.92058, -69.051611, -68.915541, -68.784529, 48.112211, 
    48.02487, 48.11558, 48.203192, 48.112211), .Dim = c(5L, 2L
    ))), class = c("XY", "POLYGON", "sfg"))), class = c("sfc_POLYGON", 
    "sfc"), precision = 0, bbox = structure(c(xmin = -69.051611, 
    ymin = 48.02487, xmax = -68.784529, ymax = 48.203192), class = "bbox"), crs = structure(list(
        input = "NAD83", wkt = "GEOGCRS[\"NAD83\",\n    DATUM[\"North American Datum 1983\",\n        ELLIPSOID[\"GRS 1980\",6378137,298.257222101,\n            LENGTHUNIT[\"metre\",1]]],\n    PRIMEM[\"Greenwich\",0,\n        ANGLEUNIT[\"degree\",0.0174532925199433]],\n    CS[ellipsoidal,2],\n        AXIS[\"latitude\",north,\n            ORDER[1],\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n        AXIS[\"longitude\",east,\n            ORDER[2],\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n    ID[\"EPSG\",4269]]"), class = "crs"), n_empty = 0L)), row.names = c(NA, 
-1L), groups = structure(list(.rows = structure(list(1L), ptype = integer(0), class = c("vctrs_list_of", 
"vctrs_vctr", "list"))), row.names = c(NA, -1L), class = c("tbl_df", 
"tbl", "data.frame")), class = c("rowwise_df", "tbl_df", "tbl", 
"data.frame"))
```

```{r, message = FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(rcoleo)
```


```{r roadmap, include = FALSE}
rcoleo:::coleo_get_required_tables() |> dplyr::filter(insectes_sol ==1)
```


```{r}
cell_data
```

Note that the column `geom` is a `POLYGON` type of object. We need to work with the literal spatial polygons that describe the shape of the cells. To get these, we need to read in the shapefiles directly, for example using `sf::read_sf("CelluleTerrain2016_2021_juin2021/CellulesTerrain2016_2021_juin2021.shp")`

### reformatting Cell data

Once we have this information, we need to reformat it to prepare for injection. We can use a function from `rcoleo` to facilitate this step

```{r cell_fmt_geom, message=FALSE}
cell_data_reformat <- cell_data |> 
  rowwise() |> 
  mutate(geom = list(rcoleo::coleo_cell_geom_fmt(geom)))

str(cell_data_reformat)
```

This step transforms the `geom` column from a spatial object into a nested list. This is what we will inject into the COLEO database.

Another important point is the use of `rowwise()` in the code above. This is a convenient way to work with dataframes going one row at a time. Here, each row contains one cell for injection. You can read more about it [here](https://dplyr.tidyverse.org/articles/rowwise.html). 

### Create the API request

Before we inject our data, we will make our requests and inspect them to make sure they work.

```{r}
cell_data_req <- cell_data_reformat |> 
  dplyr::mutate(inject_request = list(
    coleo_inject_general_df(dplyr::cur_data_all(), endpoint = "cells")
  ))
```

This step makes our request. Important things to note:

* we need to specify the endpoint where we will inject the data. Here it is `"cells"`
* we use `dplyr::cur_data_all()` to access every column of the current row of the dataset
* we use all the data in the current row. NOTE that if there is a column called "data", then `coleo_inject_general_df()` will drop it. 
* This is the step where you need to have your access key saved for Rcoleo. see the rcoleo site to set that up!
* we use `list()` to wrap the result of `coleo_inject_general_df()` because we want to store the prepared column in our dataframe, inside a list.

Let's look at this request column:

```{r, results='hold'}
print(cell_data_req$inject_request[[1]])
```

We can access the actual body of the request with:

```{r}
httr2::req_dry_run(cell_data_req$inject_request[[1]])
```

This is our last chance to make SURE we are injecting the right thing. The code at the bottom is JSON and is what will be injected into the database.

### Injecting a cell

When we are sure that we've got the right data, we can proceed to injection:

```{r}
cell_data_inject_result <- coleo_injection_execute(cell_data_req)

cell_data_inject_result
cell_data_inject_result$result
```

Here we've done two things:

* performed the request using `httr2::req_perform`. This actually injects the data
* wrapped the result in a `list()` and saved the results in a new data.frame

This second step is important -- it lets us preserve the response from the server. This response contains interesting information. For example, we can extract the ID of the cell we just injected:

```{r}
cell_data_inject_result |> 
  mutate(id = coleo_extract_id(result))
```

This is possible because when we inject data, the API sends back a response that contains the new ID that was made for the new observation.

**NOTE** the above step could also be completed manually:

```{r, eval=FALSE}
cell_data_req |> 
  dplyr::mutate(inject_result = list(httr2::req_perform(inject_request)))
```


## Injecting sites

We begin with a demonstration site:

```{r}
imaginary_sites <- tibble::tribble(
  ~cell_code,    ~site_code,    ~site_name,  ~site_type, ~site_lat,  ~site_lon, ~site_date_opened,
   "111_777",   "111_777_N09", "Mordor",     "tourbière", 48.133559,     -68.42,      "2020-01-01",
   "111_777",   "111_777_E03", "Lorien",     "forestier", 41.133559,     -67.33,      "2020-01-01"
  ) |> 
  rowwise()

```

We begin by preparing the dataset for injection. 

First, we obtain the cell ID. We need the cell ID, and not the cell_code, so we have to query the database to obtain it.

```{r}
imaginary_sites_cellID <- imaginary_sites |> 
  mutate(cell_resp = list(coleo_request_by_code(cell_code, table = "cells")),
         cell_id = coleo_extract_id(cell_resp)) 

imaginary_sites_cellID
```

this code does two things:

1. request the cell information using the `cell_code`. We do this with a convenience function called `coleo_request_by_code`. This function can take a "human readable" code and return the database record for that observation.
1. extract the code as a numeric variable.

Now that we have the `cell_id` we can drop the `cell_code` and the `cell_resp` column. We can also format the data for injection:

```{r}
imaginary_sites_for_injection <- imaginary_sites_cellID |> 
  select(-cell_code, -cell_resp) |> 
  coleo_prep_input_data(db_table = "sites")

imaginary_sites_for_injection
```

the function `coleo_prep_input_data` has automaticaly renamed the columns, reformatted the spatial information, and nested any unnecessary columns as "data".

Now we prepare the injection request, inspect them, and inject them:

```{r}
imaginary_requests <- imaginary_sites_for_injection |> 
  mutate(inject_request = list(coleo_inject_general_df(cur_data_all(), endpoint = "sites")))

# visually inspect the injection
httr2::req_dry_run(imaginary_requests$inject_request[[1]])

# run and save all the injections
imaginary_injected <- coleo_injection_execute(imaginary_requests)

imaginary_injected
```

These two sites were injected successfully!

## Campaigns

This vignette will show how to inject fake soil insect data, which will allow us to show the injection of 7 different tables:

1. campaigns
1. traps
1. landmarks
1. samples
1. observations
1. obs_species
1. ref_species

We start with some fake campaign information. This dataset contains ALL the columns we need to inject all the tables mentioned above.

```{r}
imaginary_insects <- tibble::tribble(
  ~site_code, ~campaign_type, ~campaign_date_opened, ~campaign_date_closed, ~observation_date, ~observation_is_valid, ~trap_code, ~land_lat, ~land_lon, ~sample_code, ~campaign_technician_1, ~ref_taxa_rank, ~ref_taxa_tsn, ~ref_taxa_name, ~observation_taxa_name, ~observation_variable, ~observation_value, ~observation_notes,
      "111_777_E03", "insectes_sol",          "2020-06-08",          "2020-06-30",      "2020-06-30",                  TRUE,     "A-N1",  50.56699, -74.44866,  "2020-0097",       "Legolas, Gimli",  "sous-classe",            NA, "Fake_beetleA",         "Fake_beetleA",           "abondance",                  1,                 NA,
      "111_777_E03", "insectes_sol",          "2020-06-08",          "2020-06-30",      "2020-06-30",                  TRUE,     "A-N1",  50.56699, -74.44866,  "2020-0097",       "Legolas, Gimli",       "espèce",            NA, "Fake_beetleB",         "Fake_beetleB",           "abondance",                  6,                 NA,
      "111_777_E03", "insectes_sol",          "2020-06-08",          "2020-06-30",      "2020-06-30",                  TRUE,     "A-N1",  50.56699, -74.44866,  "2020-0097",       "Legolas, Gimli",       "espèce",            NA, "Fake_beetleC",         "Fake_beetleC",           "abondance",                  10,                 NA,
      "111_777_E03", "insectes_sol",          "2020-06-08",          "2020-06-30",      "2020-06-30",                  TRUE,     "A-N1",  50.56699, -74.44866,  "2020-0097",       "Legolas, Gimli",       "espèce",            NA, "Fake_beetleR",         "Fake_beetleR",           "abondance",                  11,                 NA,
      "111_777_E03", "insectes_sol",          "2020-06-08",          "2020-06-30",      "2020-06-30",                  TRUE,    "B-NE1",  50.56698,  -74.4487,  "2020-0098",       "Legolas, Gimli",       "classe",            NA, "Fake_beetleA",         "Fake_beetleA",           "abondance",                  1,                 NA,
      "111_777_E03", "insectes_sol",          "2020-06-30",          "2020-07-21",      "2020-07-21",                  TRUE,    "B-NE1",  50.56698,  -74.4487,  "2020-0105",       "Legolas, Gimli",  "sous-classe",            NA, "Fake_beetleL",         "Fake_beetleL",           "abondance",                  3,                 NA,
      "111_777_E03", "insectes_sol",          "2020-06-08",          "2020-06-30",      "2020-06-30",                  TRUE,    "B-NE1",  50.56698,  -74.4487,  "2020-0098",       "Legolas, Gimli",       "classe",            NA, "Fake_beetleG",         "Fake_beetleG",           "abondance",                  1,                 NA,
      "111_777_E03", "insectes_sol",          "2020-06-08",          "2020-06-30",      "2020-06-30",                  TRUE,    "B-NE1",  50.56698,  -74.4487,  "2020-0098",       "Legolas, Gimli",       "espèce",            NA, "Fake_beetleH",         "Fake_beetleH",           "abondance",                  1,                 NA
  )

knitr::kable(imaginary_insects)
```


Important things to notice about the data are:

* the only site-level information is the `site_code`
* all the columns have prefixes (e.g. "*campaign_*type" or "*land_*lat). This is to differentiate columns which might have the same name but which are destined for different database tables. these will be automatically renamed during data preparation
* the format of dates is correct: AAAA-MM-JJ
* all columns have the correct values.

In a real workflow, a validation step will confirm that all of the above points are true. Getting the data in teh correct format is most of the work of injection.

### data preparation

First we need to obtain the database ID for site. The process is the same as above:

```{r get-site-id}
site_id_imaginary <- imaginary_insects |> 
  rowwise() |> 
  mutate(coleo_id = list(coleo_request_by_code(human_code = site_code, table = "sites")),
         site_id = coleo_extract_id(coleo_id)) |> 
  select(-site_code, -coleo_id) |> 
  dplyr::relocate(site_id)
```

This code queries the database for the site_id, then removes the site code.


```{r}
prepared_imaginary <- coleo_prep_input_data(site_id_imaginary, db_table = "campaigns")
```

There is one more step; we need to manually place the technicians into a list 
(NOTE should this be automated?)

```{r}
for_injection_imaginary <- prepared_imaginary |> 
  mutate(technicians = list(if_else(is.na(technician_1), true = list(NA_character_), false = list(technician_1)))) |> 
  ungroup() |> 
  select(-technician_1) |> 
  rowwise()

for_injection_imaginary

```

we're now ready to inject the data, using the same workflow as above:

```{r}
requests_imaginary <- for_injection_imaginary |> 
  mutate(inject_request = list(coleo_inject_general_df(cur_data_all(), endpoint = "campaigns")))

httr2::req_dry_run(requests_imaginary$inject_request[[1]])
```

Once we have the right format, we can perform the injections

```{r}
camp_injected_imaginary <- coleo_injection_execute(requests_imaginary)

camp_injected_imaginary
```

this is the first of *seven* tables we need to inject. Every time we inject a table, the database gives a new ID to those records. Then, we need to use thse IDs when we inject the subsequent tables. Because we have preserved the responses, this is straightforward:

```{r}

camp_id_imaginary <- camp_injected_imaginary |> 
  mutate(campaign_id = coleo_extract_id(result)) |> 
  select(campaign_id, data) |> 
  tidyr::unnest(cols = c(data))

```


## Traps

```{r}
trap_inject_imaginary <- camp_id_imaginary |> 
  coleo_prep_input_data(db_table = "traps") |> 
  mutate(inject_request = list(coleo_inject_general_df(cur_data_all(), endpoint =  "traps")))

trap_inject_imaginary
```

```{r}
trap_inject_imaginary$inject_request[[1]] |> httr2::req_dry_run()
```

```{r}
trap_response_imaginary <- trap_inject_imaginary |> coleo_injection_execute()

trap_camp_ids_imaginary <- trap_response_imaginary |>
  mutate(trap_id = coleo_extract_id(result)) |> 
  ungroup() |> 
  select(campaign_id, trap_id, data) |> 
  tidyr::unnest(cols = c(data))
```

NOTE that this time we keep both campaign and trap IDs

## landmarks

```{r}
repere_inject_imaginary <- trap_camp_ids_imaginary |> 
  coleo_prep_input_data("landmarks") |> 
  mutate(inject_request = list(coleo_inject_general_df(cur_data_all(), endpoint = "landmarks")))
```

```{r}
httr2::req_dry_run(repere_inject_imaginary$inject_request[[1]])
```

```{r}
repere_response <- coleo_injection_execute(repere_inject_imaginary)

land_id_imaginary <- repere_response |> 
  mutate(landmark_id = coleo_extract_id(result)) |> 
  ungroup() |> 
  select(ends_with("_id"), data) |> 
  tidyr::unnest(cols = c(data))
```

## Samples

```{r}
samp_inject_imaginary <- land_id_imaginary |> 
  coleo_prep_input_data("samples") |> 
  mutate(inject_request = list(coleo_inject_general_df(cur_data_all(), endpoint = "samples")))
```

```{r}
httr2::req_dry_run(samp_inject_imaginary$inject_request[[1]])

samp_response <- coleo_injection_execute(samp_inject_imaginary)

samp_response |> 
  mutate(sample_id = coleo_extract_id(result)) |> 
  ungroup() |> 
  select(ends_with("id"), data) |> 
  tidyr::unnest(cols = c(data))

```


## Observations

```{r}
coleo_prepare_injection <- function(df, db_table){
  df |> 
    coleo_prep_input_data(db_table) |> 
    dplyr::mutate(inject_request = list(coleo_inject_general_df(cur_data_all(), endpoint = db_table)))
}

coleo_polish_injection <- function(df){
  # get the name of the type of table just injected
  
  df |> 
    mutate()
}


```



### Workflow

* validate the whole dataset
* prepare the data for one table
* request IDs from the database
* if and only if the ID is absent, create a request to inject it (in a list col)
* check that these are formatted correctly by printing a few
* if and only if the ID is absent, inject that record, return response
* extract the ID code from that response
* combine ID codes from injection with those already there. Rename any columns. drop old columns. unnest the data nested in "prepare the data".



```{r, include=FALSE}
end_vignette()
```

