---
title: "telecharge_chiropteres"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{telecharge_chiropteres}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(rcoleo)

# can use get_all_observations_from_a_site to get the bats 

# first step is to download all the site information

liste_sites <- rcoleo::download_sites_sf()

liste_sites
```

this data.frame contains all the information that we need to make requests about a site. 

The convenience function `rcoleo::get_all_observations_from_a_site` will take this table as an input and request every observation from every campaign in it! 

```{r eval=FALSE}
is_what <- get_all_observations_from_a_site(liste_sites)
```

But don't do that because it takes forever! instead filter the dataset down to just what you want:

```{r}
library(tidyverse)
liste_sites %>% 
  filter(cell.name == "Forêt Montmorency") %>% 
  get_all_observations_from_a_site()
```

this gets you all the observations from Foret Montmorency

We can also filter by specific types of campaigns


```{r}
liste_sites$campaigns %>% head
```

```{r}
# let's filter out everything but the sites and campaigns we want

# if its acoustique then you should drop 

liste_sites$campaigns %>% purrr::map_lgl(~any(.$type == "acoustique"))

sites_with_target <- liste_sites %>% 
  filter(., purrr::map_lgl(.$campaigns, ~any(.$type == "acoustique")))

# and then you should filter for only those campaigns
sites_and_camps_with_target <- sites_with_target %>% 
  mutate(camp2 = campaigns %>% map(~ filter(.x, type == "acoustique")))

# now you can get em all 

# get_all_observations_from_a_site(sites_and_camps_with_target)
# except that would still be slow!!

# wrap it in a function

filter_site_df <- function(site_df, campaign_type){
  
  sites_with_target <- liste_sites %>% 
    dplyr::filter(., purrr::map_lgl(.$campaigns, ~any(.$type == campaign_type)))
  
  # and then you should filter for only those campaigns
  sites_and_camps_with_target <- sites_with_target %>% 
    dplyr::mutate(camp2 = campaigns %>% purrr::map(~ filter(.x, type == campaign_type)))

  return(sites_and_camps_with_target)
    
}
```


```{r}
acous <- query_gen("campaigns", list(type = "acoustique"))

acous %>% glimpse
liste_sites %>% glimpse

# perhaps site_code?
semi_join(liste_sites, acous, by = c("site_code" = "site.site_code"))

# seems to be the right size! 



```

